A Framework for Classification of Social Machines

### 2013-02-15 latex/sig-alternate.tex
Added more about framework, put affordances in a table

@@

Finally, we identified several contracts that are relevant for all social machines. First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time. Specifically:

\begin{enumerate}
\item Size
\item Maturity
\item Competitors
\item Diversity of user base
\item Popularity over time
\item Frequency of participation
\end{enumerate}

Next, there are constructs regarding the design affordances of the social machine:

\begin{enumerate}
\item Does it afford anonymity
\item Does it afford quality control or moderation, and in what ways
\item What types of contribution does it afford
\item Openness of resulting data set
\end{enumerate}

In the middle we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints and:

\begin{enumerate}
\item worked within them
\item worked around them
\item gamed them
\item extended and enhanced them
\item derived new structures and conventions (e.g., hashtags)
\end{enumerate}

The result is what we call `participation' constructs including:

\begin{enumerate}
\item Roles emerging in the social machine  (domain expert, troll/spammer, hub, personal gainer)
\item Quality of contributions
\item Content of contributions
\item Global / cultural reach
\item Task reach --- what do people use it for? 	
\item Active/Passive participation roles 
\end{enumerate}

>>>>

In addition to relationships to Maslow's hierarchy of needs, which drive the motivations,
we suggest that there are also resulting actions from each motivation. Thus, it should be
possible to follow basic human needs to motivations in social machines, through to the
actual actions (and therefore the necessary software feature support) that are made by
participants in social machines. Our theory is that this path may enable designers of
social machines to satisfy human needs by implemented specific features into their social
machines, or to identify potential gaps in their serving of needs and motivations due
to lack of implemented features.

@@

Finally, we identified several constructs that are relevant for all social machines. First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time. We have documented these constructs in Table \ref{table:constructs}. In the left column we list construct regarding the design affordances of the social machine. In the middle column we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints. Finally, in the right column we list analytical measures of social machines which are subject to change over time.

\begin{table*}[htbp]
\begin{center}
\begin{tabular}{|p{6.5cm}|p{5.5cm}|p{4.5cm}|}
\hline
Designed Affordances & Participatory Affordances & Analytics \\
\hline
Does it afford anonymity & worked within them & Size \\
Does it afford quality control or moderation, and in what ways & worked around them & Maturity \\
What types of contribution does it afford & gamed them & Competitors \\
Openness of resulting data set & extended and enhanced them & Diversity of user base \\
 & derived new structures and conventions (e.g., hashtags) & Population over time \\
 &  & Frequency of participation \\
\hline
\end{tabular}
\end{center}
\caption{Constructs of social machines, listing affordances that are designed into machines, affordances that have resulted from use, and analytical constructs that can be measured and change over time.} \label{table:constructs}
\end{table*}%

The result is what we call `participation' constructs. These constructs arise from use of a
social machine, and are not necessarily predictable; they include:

\begin{enumerate}
\item Roles emerging in the social machine.
   \indent In addition to designed and implemented roles such as {\it moderator} and {\it administrator}, emerging roles include motivations of the participant, such as {\it trolls}, {\it spammers}, {\it domain experts}, {\it social nexus}, and {\it re-blogger}.
\item Quality of contributions
    \indent For any social machine, the quality of contributions will vary. While some
    issues with quality can be predicted and protected against (e.g., through 
    multi-user result agreement on Mechanical Turk~\cite{ipeirotis2010quality}), some
    particular issues with quality cannot always be predicted ahead of time.
\item Content of contributions
    \indent While a system operator may design their social machine to generate specific
    types of content, participants may subvert the system in order to generate other types,
    or to concentrate on areas that were unforeseen at the time of launch, but are now seen
    as useful to the operators.
\item Global / cultural reach
    \indent 
\item Task reach --- what do people use it for?
\item Active/Passive participation roles 
\end{enumerate}
